### 2020-08-26 analysis.Rmd
switched to my old data from esdlflume, updated with Candace's runs.

---
title: "Particle Interception as a Function of Reynolds Number and Collector Density"
author: "Jordan Wingenroth"
date: "8/26/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

## Objective(s)

In the interest of documenting our analytic methods, I've decided to make one comprehensive, hopefully lay-readable document including all steps.

## Data sourcing

We entered data into Google Sheets as we processed samples in the lab. I downloaded .csv versions of these data in 2019, then made a few modifications as I explored confounders. Today, I added data from runs done in my absence to this dataset. Field names, date syntax, and other conventions weren't kept consistent, so some data wrangling is necessary.

```{r, message=FALSE, warning=FALSE}

library(tidyverse)

```

```{r}

pumpfiles <- list.files("data/peristaltic pumps/")
trapfiles <- list.files("data/sediment traps/")

length(pumpfiles)
length(trapfiles)

```