### 2020-08-27 analysis.Rmd
finished pump data part of analysis.rmd

---
title: "Particle Interception as a Function of Reynolds Number and Collector Density"
author: "Jordan Wingenroth"
date: "8/26/2020"
toc: TRUE
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

## Objective(s)

In the interest of documenting our analytic methods, I've decided to make one comprehensive, hopefully lay-readable document including all steps.

## Data sourcing and tidying

We entered data into Google Sheets as we processed samples in the lab. I downloaded .csv versions of these data in 2019, then made a few modifications as I explored confounders. Today, I added data from runs done in my absence to this dataset. Field names, date syntax, and other conventions weren't kept consistent, so some data wrangling is necessary. I also removed the blank data at the beginning of some files manually.

```{r, message=FALSE, warning=FALSE}

library(tidyverse)

```

```{r}

pumpfiles <- list.files("data/peristaltic pumps/")
trapfiles <- list.files("data/sediment traps/")

length(pumpfiles)
length(trapfiles)

```

As you see, there are different numbers of files for sediment trap versus peristaltic pump data.

```{r}

pumpdate <- str_sub(pumpfiles,0,6)
trapdate <- str_sub(trapfiles,0,6)

tibble(pumpdate, pumpdate %in% trapdate)

```

In all, five runs are missing sediment trap data, meaning they won't be used in our final, published analyses.

### Suspended sediment data

Let's turn our attention to the suspended concentration data (i.e., pump data).

```{r, message=FALSE, warning=FALSE}

pumpdata <- lapply(pumpfiles, function(x) read_csv(paste0("data/peristaltic pumps/",x)))

names(pumpdata) <- pumpdate

x <- pumpdata

tidypump <- lapply(seq_along(x), function(i) {
  select(x[[i]], 
         loc = Location, 
         ht = Height,
         t = `time series`,
         mvc = contains("(ppm)")) %>%
    filter(t < 21) %>% #filter a few timepoints outside the normal window 
    mutate(t = (t-min(t)+1)*300, #convert from timestep to seconds
           mvc = as.numeric(mvc), 
           date = as.numeric(names(x)[[i]])) %>%
    filter(mvc < 80) #filter out a few early-timestep anomalies
  }
  )

pump <- bind_rows(tidypump)

pump

```

So now we have a nice long table of our pump data (time x concentration) stratified by run (i.e., treatment), height, and upstream/downstream location.

Let's plot the decay curves separately to see what the data looks like qualitatively.

```{r}

pump %>%
  ggplot(aes(x = t, y = mvc)) +
  geom_jitter(alpha = .5, height = 0) +
  facet_wrap(~date)

```

Aside from a few runs where **part** or **all** of the data don't follow the decay pattern, the results look pretty good! If we repeated runs as necessary to replace those with erroneous data, that shouldn't be a problem. Here I'll narrow down our dataset by knocking out runs with clear issues.

```{r}

pumpfinal <- pump %>%

#first 3 runs had starting sediment mass of 100g rather than 200g

    filter(date > 181005) %>%

#we can't use runs without sediment mass

    filter(as.character(date) %in% trapdate)

length(unique(pumpfinal$date))

```

So this leaves 20 rows. Let's join this to our run metadata table.

```{r}

metadata <- read_csv("data/run_metadata.csv")

sum(unique(pumpfinal$date) %in% metadata$date)

```

Hoorah! We have metadata for all of our runs of interest.

```{r}

pumpfinal <- left_join(pumpfinal, metadata, by = "date")

```

Now then, let's remove the biofouled runs, which won't be used in assessing our primary hypothesis about Reynolds number and collector density, and we'll see what we're left with.

```{r}

pumpfinal %>%
  filter(growth_days==0) %>%
  ggplot(aes(x = t, y = mvc)) +
  geom_jitter(alpha = .5, height = 0) +
  facet_wrap(~date)

```

These runs do look quite familiar to me because I think this is about my 10th time through this analysis. `190417` is the run with clear issues, in the early timesteps more specifically. During one of those previous analyses, I found that there were several runs where **part** of the timespan was clearly out of line with the expected decay pattern. Fortunately we tracked metadata about the logistics of laboratory analysis, and it just so happens that the clearly erroneous data were all from samples processed by one researcher in particular. `190417` is the only run out of these 13 where they were involved in processing samples, at least since we started recording initials on our work. We repeated that run on `190506` and got results more in line with exponential decay.

So that leaves 13 runs. Aside from one run at 232 collectors/$\text m^2$, the result of a miscalculation during dowel installation, the 12 remaining are a perfect match for our parameter space of 4 densities (including 0 control) and 3 velocities/Re values.

```{r}

pumpfinal %>%
  filter(growth_days==0, date != 190417, dowel_density != "0232") %>%
  ggplot(aes(x = t, y = mvc)) +
  geom_jitter(alpha = .5, height = 0) +
  facet_grid(pump_freq~dowel_density)

```